---
title: "Question1"
author: "Cesar Conejo"
date: "3/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r}
temps_7 <- read.table(file = "data/temps-7.txt", header = TRUE)
temps_other <- read.table(file = "data/temps-other.txt", header = TRUE)

temps_7 <- temps_7$x
temps_other <- temps_other$x
```


```{r}
summary(temps_7)
summary(temps_other)
```

## a. Perform a kernel density estimation for temps-7 and temps-other using what you consider is the most adequate bandwidth. Since the temperatures are positive, is it required to perform any transformation?

```{r}
par(mfrow = c(1,2))
hist(temps_7, main = 'Problematic battery')
hist(temps_other, main = 'Past battery')
```


```{r}
par(mfrow = c(1,2))
plot(density(x = temps_7), xlim = c(-5,65), main = 'Problematic battery')
plot(density(x = temps_other), xlim = c(-5,65), main = 'Past battery')
```

```{r}
# Deal with the problem of Select a particular bandwidth
par(mfrow = c(1,2))
plot(density(x = temps_7))
plot(density(x = temps_7, bw = 0.5))
```


```{r}
# Transformed the data: No necessary. The min observation is away from zero! (No assigment of probability)

# kde with log-transformed data
kde <- density(log(temps_7))
plot(kde, main = "Kde of transformed data")
range(kde$x)

# Untransform kde$x so the grid is in (0, infty)
kde_transf <- kde
kde_transf$x <- exp(kde_transf$x)
# Transform the density using the chain rule
kde_transf$y <- kde_transf$y * 1 / kde_transf$x

par(mfrow = c(1,2))
plot(density(x = temps_7), xlim = c(-5,65), main = 'Problematic battery')
# Transformed kde
plot(kde_transf, main = "Problematic battery: Transformed kde", xlim = c(-5, 65))
```


```{r}
## SAME WITH temps_other

## kde with log-transformed data
kde <- density(log(temps_other))
plot(kde, main = "Kde of transformed data")
range(kde$x)

# Untransform kde$x so the grid is in (0, infty)
kde_transf <- kde
kde_transf$x <- exp(kde_transf$x)
# Transform the density using the chain rule
kde_transf$y <- kde_transf$y * 1 / kde_transf$x

par(mfrow = c(1,2))
plot(density(x = temps_other), xlim = c(-5,60), main = 'Past battery')
# Transformed kde
plot(kde_transf, main = "Past battery: Transformed kde", xlim = c(-5, 60))
```

## b. Is there any important difference on the results from considering the LSCV selector over the DPI selector?

### temps_7

```{r}
summary(temps_7)
par(mfrow = c(1,2))
plot(density(x = temps_7))
plot(density(x = temps_7), xlim = c(-5,65))
```


```{r}
bw.nrd0(x = temps_7) # Default option!!!


bw.nrd(x = temps_7)  # Rule-Thumb 

bw.ucv(x = temps_7)

# Eduardo function

source("functions/bw_ucv_mod.R")
bw_ucv <- bw.ucv.mod(x = temps_7, plot_cv = TRUE, h_grid = 10^seq(-1.5, 0.5, l = 200))
bw_ucv

bw_dpi <- bw.SJ(x = temps_7, method = "dpi")
ks::hpi(temps_7)
```

```{r}
plot(density(x = temps_7), xlim = c(-5,65))
```

```{r}
plot(density(x = temps_7, bw = bw_ucv), xlim = c(-5,65))
```

```{r}
plot(density(x = temps_7, bw = bw_dpi), xlim = c(-5,65))
```


## c. It seems that in temps-7 there is a secondary mode. Compute a kernel derivative estimation for temps-7 and temps-other using what you consider are the most adequate bandwidths.

```{r}
plot(density(x = temps_7), xlim = c(-5,65))
pos_Max <- which.max(density(x = temps_7)$y)
pos_Max
density(x = temps_7)$x[pos_Max]

plot(density(x = temps_7), xlim = c(-5,65))
abline(v = density(x = temps_7)$x[pos_Max] )


## Find the minimun (Then, from the minumn, look up the max in the rest of the dist)

# MIN
YY <- density(x = temps_7)$y[density(x = temps_7)$x > 20 & density(x = temps_7)$x < 40]
minY <- min(YY)
posMin <- which(density(x = temps_7)$y == minY)
density(x = temps_7)$x[posMin]
## [1] 33.83911

# MAX
YY <- density(x = temps_7)$y[density(x = temps_7)$x > 33 & density(x = temps_7)$x < 60]
maxY <- max(YY)
posMax <- which(density(x = temps_7)$y == maxY)
density(x = temps_7)$x[posMax]
```


```{r}
plot(density(x = temps_7), xlim = c(-5,65), main = "temp_7: Density Estimation")
abline(v = density(x = temps_7)$x[pos_Max], col = "3")
abline(v = density(x = temps_7)$x[posMax], col = "3")
```

```{r}
kdde_0 <- ks::kdde(x = temps_7, deriv.order = 0)
plot(kdde_0, xlab = "x", main = "Density estimation")
abline(v = density(x = temps_7)$x[pos_Max], col = "3")
abline(v = density(x = temps_7)$x[posMax], col = "3")
```


```{r}
# Density derivative estimation (automatically chosen bandwidth, but different
# from kdde_0!)
kdde_1 <- ks::kdde(x = temps_7, deriv.order = 1)
plot(kdde_1, xlab = "x", main = "Density derivative estimation")
abline(v = density(x = temps_7)$x[pos_Max], col = "3")
abline(v = density(x = temps_7)$x[posMax], col = "3")
abline(h = 0)
```


```{r}
kdde_2 <- ks::kdde(x = temps_7, deriv.order = 2)
plot(kdde_2, xlab = "x", main = "Density second derivative estimation")
abline(v = density(x = temps_7)$x[pos_Max], col = "3")
abline(v = density(x = temps_7)$x[posMax], col = "3")
abline(h = 0)

```


 $\hat{h}_{LSCV}$